Computational methods in landscape ecology

### 2024-09-17 Manuscript.qmd
review of the discussion by jn

affiliations: 
  - id: 1
    name: International Institute for Applied Systems Analysis, Biodiversity, Ecology, and Conservation Group, Laxenburg, Austria
  - id: 2
    name: Adam Mickiewicz University, Poznan, Poland
  - id: 3
    name: University of Münster, Münster, Germany
  - id: 4
    name: École polytechnique fédérale de Lausanne, Lausanne, Switzerland
  - id: 5
    name: Carl R. Woese Institute for Genomic Biology, University of Illinois at Urbana-Champaign, Urbana, IL, USA

@@

papersize: a4paper
fontsize: 12pt
geometry: margin=25mm
indent: 2m
linestretch: 1.5
number-sections: true

>>>>

affiliations: 
  - id: 1
    name: International Institute for Applied Systems Analysis, Biodiversity, Ecology, and Conservation Group, Laxenburg, Austria
  - id: 2
    name: Adam Mickiewicz University, Poznań, Poland
  - id: 3
    name: University of Münster, Münster, Germany
  - id: 4
    name: École polytechnique fédérale de Lausanne, Lausanne, Switzerland
  - id: 5
    name: Carl R. Woese Institute for Genomic Biology, University of Illinois at Urbana-Champaign, Urbana, IL, USA

@@

papersize: a4paper
fontsize: 10pt
geometry: margin=25mm
indent: 2m
linestretch: 1.5
number-sections: true

### 2024-09-17 Sections/08-discussion.qmd
review of the discussion by jn

Since it emerged in the 1980s [@Turner2005LandscapeEcology], the field of landscape ecology has constantly evolved [@Costanza2019DescribingAnalyzing; @Gustafson2019HowHas] and is now highly depending on computational methods.
Thus, future progress within the research field will largely depend on available software and analysis tools [@Costanza2019DescribingAnalyzing].
To maximize the benefits of "Open science", such as shareability, reproducibility, and transparency, the research community will gain the most if both existing and new software are made available under open-source licenses and support script-based workflows [@Powers2019OpenScience; @Vogt2019PatternsSoftware].

Furthermore, landscape ecology will benefit and adapt solutions from technological advancements in other fields.
This includes advancements in remote sensing technology, with increased data availability and diversity from a range of sources, including large satellites, unmanned aerial vehicles, and ground-based sensor systems. 
These systems provide both passive sensors such as multispectral, hyperspectral, and thermal sensors, and active optical sensors such as LiDAR [@Crowley2020RemoteSensings].
In combination with other increasingly available data sources, e.g, citizen science data, ecology enters the era of big data requiring methods that can handle heterogeneous data or high-throughput computing resources [@Farley2018SituatingEcology].
Due to its high flexibility and performance artificial intelligence has become popular and deep learning and machine learning algorithms can be used for e.g., mapping, classification and feature extraction, modeling, or prediction and simulations [@Christin2019ApplicationsDeep; @Stupariu2022MachineLearning; @Pichler2023MachineLearning].
Quantum computing offers a potential pathway to surpass the limits of current computational technologies [@Woolnough2023QuantumComputing].

@@

>>>>

Since emerging in the 1980s [@Turner2005LandscapeEcology], the field of landscape ecology has constantly evolved [@Costanza2019DescribingAnalyzing; @Gustafson2019HowHas] and is now highly depending on computational methods.
Thus, future progress within the research field will largely depend on available software and analysis tools [@Costanza2019DescribingAnalyzing].
To maximize the benefits of "Open science", such as shareability, reproducibility, and transparency, the research community will gain the most if both existing and new software support script-based workflows and are made available under open-source licenses [@Powers2019OpenScience; @Vogt2019PatternsSoftware].
<!--JN: maybe add a sentence or two summarizing the previous sections (e.g., that this paper presents various computational methods used in landscape ecology)-->

<!-- JN: the next sentence seems to be broken -->
Furthermore, landscape ecology will benefit and adapt solutions from technological advancements in other fields.
This includes advancements in remote sensing technology, with increased data availability and diversity from a range of sources, including large satellites, unmanned aerial vehicles, and ground-based sensor systems. 
These systems provide both passive sensors such as multispectral, hyperspectral, and thermal sensors, and active optical sensors such as LiDAR and SAR[@Crowley2020RemoteSensings].
In combination with other increasingly available data sources, e.g, citizen science data, ecology is in the era of increasingly large data requiring methods that can handle heterogeneous data or high-throughput computing resources [@Farley2018SituatingEcology].
Due to its high flexibility and performance, artificial intelligence has become popular and deep learning and machine learning algorithms can be used for e.g., mapping, classification and feature extraction, modeling, or prediction and simulations [@Christin2019ApplicationsDeep; @Stupariu2022MachineLearning; @Pichler2023MachineLearning].
Quantum computing may additionally offer a potential pathway to surpass the limits of current computational technologies [@Woolnough2023QuantumComputing].

@@

<!-- JN: additional possible topics to discuss here is that the computational methods are also related to the developments in computer science; thus, new frameworks, tools, good practices, etc. are constantly emerging, and not everything is about new methods, but also about new ways to implement them, new ways to share them, better ways to document them, more consistent syntaxes, etc. -->
